The Host Galaxies of 106 Rapidly Evolving Transients discovered by the Dark Energy Survey

### 2020-07-31 referee_response_1.tex

\item I also think the second sentence is misleading, that RETs are characterized by light curves that decline much faster than common classes of supernovae and can be seen to redshifts greater than 1. Again, there is considerable heterogeneity in what has been called a RET/FBOT in the literature. In Drout +14, the distinction was that the peak of the light curve could not be powered by nickel decay — but a number of events had decline rates that could easily be explained by nickel decay (see Section 8.1). In SN2018gep, the rise and decline rate would have passed the Drout+14 criteria, but the decline was consistent with a Type Ic-BL SN. 18cow, of course, did decline much more quickly than typical SNe. Regarding the “can be seen to redshifts greater than 1,” this is only true of the very luminous RETs. The sample in Drout+14 spanned an enormous range of peak luminosity. I would remove this sentence, and instead make the point that the main distinguishing feature of an RET is that the *peak* of the light curve is unlikely to be powered by the radioactive decay of Ni-56 — and that the light curve is single peaked, unlike the many double peaked SNe that are known. In general, I think that it is typically the *rise* time that distinguishes RETs, rather than the decline timescale. Perhaps give numbers for the range of right times, decline times, and peak luminosities that are observed for these events?

\vskip0.1cm
{\bf RESPONSE: } These two points have been addressed together, by streamlining the introductory part of the abstract to the following: ``\textit{RETs typically rise to peak in less than 10 days and fade within 30, a timescale unlikely to be compatible with the decay of Nickel-56 that drives conventional SNe. Their peak luminosity spans a range of $-15<M_g<-22.5$, with some events observed at redshifts greater than 1.}"

\item There should be an “of” between “common classes” and “supernovae”
\vskip0.1cm
{\bf RESPONSE: } Changed to ``conventional supernovae"

\item I would state the exact definition of a DES RET somewhere in the abstract
\vskip0.1cm
{\bf RESPONSE: } We appreciate the desire to properly define RET in the abstract; however there is no exact definition of a DES RET that does not take up a full section, as per Section 2.2. Instead we have added the typical rise and decline times.

\item In the abstract, perhaps state what is meant by “metallicity” ([O/H], etc)?
\vskip0.1cm
{\bf RESPONSE: } We have added ([O/H])

@@

\item “We find that RETs prefer galaxies with high specific SFRs” — state the number or range
\vskip0.1cm
{\bf RESPONSE: } Added the mean

\item “appear to show a lack of chemical enrichment” — state the number or range
\vskip0.1cm
{\bf RESPONSE: } Added the mean

\item “no clear relationships between properties of the host galaxies” — what properties?
\vskip0.1cm
{\bf RESPONSE: } Added mass and sSFR

\item Perhaps include a final sentence looking to the future? What is the next step?
\vskip0.1cm
{\bf RESPONSE: } Such a sentence was originally in the abstract, but was cut due to the MNRAS abstract word limit. We provide an outlook to future work at the end of the paper.
\end{enumerate}

@@

\item[] Table 1: Perhaps add “decimal degrees” to the RA and Dec columns?
\vskip0.1cm
{\bf RESPONSE: } Added `deg' 
\item[] Figure 2: Please indicate the orientation (i.e. a compass showing N/S, E/W). I also suggest putting a scale bar on the figure rather than just stating it in the caption. Consider labeling the relevant boxes with the host redshift as well.

{\bf RESPONSE: } Added all of the above.

\end{enumerate}
\section*{Section 2.3}
\begin{enumerate}
    \item “there is no other large sample of RETs with which to compare host galaxy properties.” Do the properties of the DES RET sample overlap with known classes of supernovae? For example, Type Ibn supernovae, which rise quite quickly to high peak luminosity (M ~ -19.5)? It might be useful to acknowledge that there are probably subclasses represented in the DES sample, which arise from a variety of progenitor channels.

    \vskip0.1cm
    {\bf RESPONSE: } We have added a sentence to the beginning of 2.3.2: \textit{It is likely that the progenitor and explosion scenario of the DES RET sample is heterogeneous in origin and may include some known classes of transient, and as such we choose samples of host galaxies of various transients from the literature to compare the DES RET hosts to}.

\item The reference to SN2018gep here, Ho et al. (2019), is actually to the AT2018cow paper. Please add the correct SN2018gep reference, which is a separate paper, i.e. will end up being Ho et al. 2019a \& Ho et al. 2019b.
\vskip0.1cm
 {\bf RESPONSE: } Thanks! Changed.

 \item “with the the” — extra “the”
 \vskip0.1cm
 {\bf RESPONSE: } Thanks, removed

 \item I might suggest adding iPTF16asu to your list of comparison objects. It was similar to SN2018gep. The reference is Whitesides+ 2017
 \vskip0.1cm
 {\bf RESPONSE: } Thanks for pointing this out. We have added iPTF16asu to the comparison objects.

@@

\end{enumerate}

\section*{Section 2.4}

\begin{enumerate}
    \item I was a bit surprised by the choice of SDSS as a comparison sample of field galaxies. My understanding is that SDSS galaxies suffer from a number of biases -- they will be highly incomplete for dwarfs and biased towards massive galaxies, and the metallicity measurements are very biased (since they have enough S/N in the critical lines for the measurement, and the strengths of those lines is metallicity-dependent.) Since RETs are likely related to GRBs and SLSNe in terms of their stellar evolutionary pathways, it seems important to have a control sample that is inclusive of low-mass galaxies (see the discussion in Section 4.2 in Modjaz+2019, which presents a sample of Ic and Ic-BL hosts from PTF). Why do the authors not use a volume-complete sample of local galaxies instead or as well, as was done in e.g. Modjaz+2019? Either way, I think the authors need to spend more time motivating their choice of field galaxy sample, and discussing possible drawbacks and biases.
   \vskip0.1cm
 {\bf RESPONSE: } We agree that SDSS is not an ideal comparison sample due to the biases pointed out by the referee. We note, however, that the DES host galaxy photometry is similarly incomplete, albeit down to lower mass/higher redshift than SDSS. Furthermore, the majority of the hosts were followed up with OzDES with a magnitude-based selection and limit. We therefore expect similar biases in the DES RET hosts as seen in SDSS. 
 We acknowledge the use of comparing to the Local Volume Legacy Survey galaxies and have included that in the figures, but this is mainly for illustrative purposes, since the bulk of low mass LVL galaxies would not be seen in DES-SN at the redshifts of the RET sample.

@@

\end{enumerate}

\section*{Section 4.2}
1) “There are several reasons this measurement may differ from one made at the SN explosion site” — could the authors give a quantitative sense for by how much this could differ?
   \vskip0.1cm
 {\bf RESPONSE: } We have added a sentence describing how these difference are negligible compared to the uncertainty caused by low SNR.

\section*{Section 5}
1) Again, I might suggest including iPTF16asu (Whitesides+2017)
 \vskip0.1cm
 {\bf RESPONSE: } Done, see above

@@

 \section*{Section 5.2}

 1)  “we compare the chemical state of RET host galaxies with...star-forming field galaxies.” As I stated
above, SDSS metallicities are very biased. I think the authors need to discuss this, or possibly use an additional comparison sample.
\vskip0.1cm
 {\bf RESPONSE: } Done, as discussed previously

 \section*{Section 6.2}
 \begin{enumerate}
     \item Ho et al. (2020) (ZTF18abvkwla) and Coppejans (2020) (CSS161010) also performed rate estimates of RETs, so these should be referenced
    \vskip0.1cm
 {\bf RESPONSE: }  We have added these references
 \item As stated in the point above, I think that for a rate estimate it is important to be clear how these different papers define RET...Drout+14 spans a large luminosity distribution, for example
\vskip0.1cm
 {\bf RESPONSE: } We have added this caveat

### 2020-07-31 DES_RET_Hosts_replyref.tex

% Abstract of the paper
\begin{abstract}
Rapidly evolving transients (RETs), also termed fast blue optical transients (FBOTs), are a recently-discovered group of astrophysical events that display rapid luminosity evolution. RETs typically rise to peak in less than 10 days and fade within 30, a timescale unlikely to be compatible with the decay of Nickel-56 that drives conventional SNe. Their peak luminosity spans a range of $-15 <M_g<-22.5$, with some events observed at redshifts greater than 1. Their evolution on fast timescales has hindered high quality follow-up observations, and thus their origin and explosion/emission mechanism remains unexplained. In this paper we define the largest sample of RETs to date, comprising 106 objects from the Dark Energy Survey, and perform the most comprehensive analysis of RET host galaxies. Using deep-stacked photometry and emission-lines from OzDES spectroscopy, we derive stellar masses and star-formation rates (SFRs) for 49 host galaxies, and metallicities ([O/H]) for 37. We find that RETs explode exclusively in star-forming galaxies and are thus likely associated with massive stars. Comparing RET hosts to samples of host galaxies of other explosive transients as well as field galaxies, we find that RETs prefer galaxies with high specific SFRs ($\langle\log \left(\mathrm{sSFR}\right)\rangle \sim -9.6$ ), indicating a link to young stellar populations, similar to stripped-envelope SNe. RET hosts appear to show a lack of chemical enrichment, their metallicities akin to long duration gamma-ray bursts and superluminous SN host galaxies $\langle12+\log\left(\mathrm{O/H}\right) \rangle \sim 9.4$. There are no clear relationships between \replyref{mass or SFR} of the host galaxies and the peak magnitudes or decline rates of the transients themselves.

@@

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

@@

There are a small number of other objects in the literature that would pass the DES RET criteria, including SNLS04D4ec \citep{Arcavi2016} at $z\sim0.59$, ZTF18abvkwla \replyref{whose rapid evolution has been attributed to shock interaction, while radio emission can be explained by a collimated jet \citep{Ho2020} at $z\sim0.27$, and the rapidly-evolving SN Ic-BL-like iPTF16asu \citep{Whitesides2017} at $z\sim0.19$. }

There are also a limited number of events detected in the local Universe whose properties appear consistent with the RETs seen at cosmological distances in the PS1 and DES samples, the most widely studied of which is AT2018cow \citep[e.g.][]{Prentice2018,Margutti2019,Perley2019}. \replyref{AT2018cow has drawn many comparisons to the cosmological RETs from DES and PS1 \citep[e.g.][]{Perley2019,Margutti2019,Fox2019,Mohan2020} due to its rapid evolution and blue colour. AT2018cow declined from its discovery, with constraints on the rise time of 1 day, and from X-rays through to radio wavelengths did not resemble any known SN, GRB afterglow, or kilonova (KN; \citealt{Ho2019}). It displayed a contracting photosphere as well as evidence for central-engine power alongside an unusual spectrum that showed similarities to broad-lined SNe Ic (SN Ic-bl) at early stages \citep[e.g.][]{Xu2018,Izzo2018}, developing to something entirely different at later epochs \citep{Perley2019} with hints of similarities to interacting SNe Ibn \citep{Fox2019}.} There are many diverse explanations for the power source of AT2018cow \replyref{proposed} in the literature, including: \replyref{engine-driven stellar explosions \citep{Ho2019,Margutti2019,Mohan2020}}; electron capture collapse of merged white dwarfs \citep{Lyutikov2019}; a tidal disruption event (TDE) of a white dwarf \citep{Kuin2019} or of a main sequence star by an intermediate mass black hole \citep{Perley2019}; common envelope jets supernova (CEJSN; \citealt{Soker2019}); or a wind-driven transient \citep{Piro2020,Uno2020}.
Other nearby \replyref{($z<0.1$)} rapid transients \replyref{that would pass the DES search criteria} include the local fast-declining SN-like transient SN2018kzr \citep{McBrien2019} which is explained by the accretion-induced collapse of a white dwarf or a white dwarf-neutron star\replyref{black hole} merger \citep{Gillanders2020}, KSN-2015K \citep{Rest2018} whose fast rise and decline is explained by the shock of an SN running into previously-expelled material, \replyref{SN2018gep \citep{Ho2019a} \replyref{which is a rapidly-declining, spectroscopically classified, nearby SN Ic-BL with a rapid rise to peak}, and the H-rich, relativistic CSS161010 \citep{Coppejans2020}.} \replyref{These objects evolved rapidly enough to have passed the DES RET criteria (Section \ref{subsec:new_method}) even at $z\sim1$ where they would appear slower, although they (or their host galaxies) may not necessarily have been bright enough to be detected.}

@@

The original method of finding RETs in the DES-SN data (and presented in \citetalias{Pursiainen2018}) was designed to be simple and used light curve modelling with Gaussian and linear fits. The simplistic method made it possible to look for exotic transients without knowing their observed characteristics beforehand and resulted in a large sample of photometrically selected fast transients. However, as the search was simplistic and relied heavily on visual inspection of the available data (etc. images, light curves, host galaxy information), it is impossible to quantify the completeness of the sample. For instance, due to the large redshift range within the sample it is entirely possible that distant events with longer rise times could have been missed due to time dilation stretching their lightcurves, while faster events \replyref{at} low redshift may evolve on timescales quicker than the survey cadence. Here, a more sophisticated search method is presented. As only a fraction of transients in DES-SN have redshift information from their host galaxies, we perform the search in the observer frame. The key features of RETs that separate them from most traditional SNe types are the fast light curve evolution \replyref{($t_{\mathrm{half}}\leq12$d)} and blue colour at peak ($g-r \lesssim 0.5$). Even though both of these quantities depend on the redshift of the transient \replyref{(timescales, for example, are dilated by a factor of 2 at $z=1$ while higher-redshift objects may appear bluer, since the hot and bright UV is shifted into the observed bands)}, they still effectively distinguish the fast events from traditional SN types. \replyref{Here, we} attempt to select a sample based on observed rise times and colours at peak brightness.

@@

In order to apply selection criterion i) from Section \ref{subsubsec:trise_gr} it is necessary to distinguish between DES-SN candidates that are truly multi-season events (typically active galactic nuclei; AGN) and those that are single-season events with spurious detections in other seasons. 
To detect and filter out AGN we use a basic convolutional neural network (CNN) classifier. \replyref{The CNN uses features of the GP light curves in a set of layers in order to ``learn" the difference between SN and AGN lightcurves. We implement two 1-dimensional feature layers: the first searches for evolution over 2-day timescales in all four filters, while the second deals with long-term evolution on 20-day timescales. We include a dropout layer after each filter layer to prevent over-fitting, and a max-pooling layer to increase speed. At the end, a softmax layer outputs a probability that the individual transient is SN-like, for which we use 50\% as the threshold}.  We train the CNN on spectroscopically confirmed SNe of all types as well as the 92 RETs identified using the \citetalias{Pursiainen2018} method (181 objects) and on spectroscopically typed AGN (182 objects), and use it to separate the sample into two photometric subtypes: AGN-like and SN-like. The classifier returns SNe-like objects with an accuracy of 0.992 on the test set (391 SNe and 79 AGN). No SNe were classified as AGN-like. The remaining AGN classified as SNe-like are removed by manual vetting later in the process. The CNN does not separate the SN-like objects into RETs and other SN subtypes: this is done in the subsequent processing step.

@@

To show how RETs compare to the galaxy population as a whole, we use a sample of $\sim800,000$ measurements from the MPA-JHU catalogues of stellar masses (based on the methods of \citealt{Kauffmann2003,Salim2007}), SFRs (based off \citealt{Brinchmann2004}), and metallicities (based off \citealt{Tremonti2004}) from the  catalogues of SDSS Data Relase 7 (DR7; \citealt{Abazajian2009}). The mean redshift is 0.08 which is much lower than for the RETs, such that significant evolution in the galaxy population has happened between the majority of RET hosts and the SDSS sample. We do not correct for this, but take it into account when analysing our findings. \replyref{The magnitude-limited nature of the SDSS redshift catalogue means that it is biased against low-mass galaxies, and metallicity estimates requiring strong lines to pass SNR thresholds mean that they too will be biased. Many works exploring extreme transients such as LGRBs, GRB-SNe, and SLSNe, use a complete, volume limited sample of field galaxies such as the Local Volume Legacy Survey (LVL) sample as a comparison \citep[e.g.][]{Chen2017,Modjaz2019,Taggart2019}. However the spectroscopic follow-up of DES RETs was performed as part of the magnitude-limited OzDES survey (Section \ref{subsec:spec}), which has similar selection biases and incompleteness effects to SDSS, albeit shifted to a higher redshift. In comparison, the many dwarf galaxies in LVL would not be detected in DES photometry, or get a redshift measurement from OzDES, and thus such a comparison sample is not representative of the way the DES RET host sample was observed. For completeness and illustrative purposes only, we display both the SDSS and LVL comparison samples.}

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{figs/RET_BPT.png}
\caption{Baldwin-Phillips-Terlovich (BPT) diagram for RET hosts, showing that the emission lines are consistent with being generated by star formation rather than AGN activity. The three curves show the delimitation between star formation and AGN at $z =0, 0.5, 1$ according to \citet{Kewley2013}. The redshift of each RET is indicated by the colourbar on the right.
\label{fig:bpt}}
\end{figure}

@@

Due to the low S/N of the spectra in this sample, we are constrained to a subset of metallicity diagnostics by the availability of only a handful of the strongest emission lines, namely \halpha, \hbeta, \OII 3727, \OIII 4959/5007, \NII 6548/6583, and \SII 6717/6731. Furthermore, for each host galaxy only a subset of these lines is detected - for example, \halpha, \NII~and \SII~are redshifted out of the spectral coverage at $z>0.3$, leaving only the oxygen and \hbeta~lines available and thus the R23 diagnostic, which is based off the relative strengths of the \OII and \OIII lines. For hosts at $z<0.3$ we are able to use the \OIII /\NII~(O3N2), \NII /\halpha~(N2), and \SII /\NII~(S2N2) line ratios.

Due to the redshift range of our sample, and the limited wavelength coverage of the spectra ($3000-8000$~\AA), we are unable to use a single line ratio to estimate the oxygen abundances. We thus determine a set of indicators for which to calculate abundances. For the O3N2 and N2 indicators we use the calibration of \citet{Pettini2004} (PP04), and if \SII~is detected we derive an abundance using the S2N2 diagnostic of \citet{Dopita2016} (D16). For the R23 indicator, we use the calibration of \citet{Kobulnicky2004} (KK04). At abundances around $12 + \log \mathrm{(O/H)} \sim 8.4$, the R23 indicator becomes two-tailed, with a low and a high value of metallicity corresponding to a single R23 ratio. In cases where the lines are available, we break this degeneracy by cross-calibrating with the \NII/\OII~ratio \citep{Kewley2008}. In the cases where \NII~is not available, and there are no other diagnostics that can be used to inform the choice of branch, we use the host galaxy stellar mass to derive a crude metallicity estimate from the mass-metallicity relation (MZR) of \citet{Kewley2008} based upon the PP04 O3N2 diagnostic. For $12 + \log \mathrm{(O/H)}_{\mathrm{MZR}} < 8.4$ we chose the lower branch, while for higher MZR metallicities we choose the upper branch. We note that this is a rough estimation. If we leave the branch choice for those with no \NII~to be random, we find that the results are consistent to within uncertainties.

The samples to which we compare metallicities span different redshift ranges, were observed with different equipment, and in many cases were compiled before certain (particularly the D16) diagnostics were devised. Therefore, in order to compare oxygen abundances between different samples we transform all abundances onto the PP04 O3N2 scale using the conversion factors given in \citet{Kewley2008}. This is not possible for the D16 diagnostic, so we discard it from the rest of our analysis, although for completeness we provide it for DES RET hosts where available. For samples that quoted multiple diagnostics, or for which sufficient line flux measurements were provided from which to calculate multiple diagnostics, we transform them all to the PP04 O3N2 scale. Following the prescription of \citet{Kruehler2015} we simultaneously minimise the oxygen abundance against the PDFs of the various different diagnostics scaled to PP04 O3N2, resulting in a final `best' PDF. We take $1\sigma$ uncertainties from the 16\textsuperscript{th} and 84\textsuperscript{th} percentiles of these PDFs, and for DES RETs display the results in Table \ref{tab:derived}.
\begin{figure}

\includegraphics[width=0.5\textwidth]{figs/SFR_Mike_RETs.png}
\caption{The $M_* -$ SFR sequence of RET hosts, hosts of other rapid transients and SNe II, as well as SDSS field galaxies (grey contours) and \replyref{LVL galaxies (grey circles)}. SFRs have not been corrected for redshift evolution. RET hosts lie slightly above the low-$z$ star-formation main sequence upper half of the SDSS contours, and systematically avoid passive galaxies (dense SDSS contours in the lower right).
\label{fig:sfms_sfr}}
\end{figure}

@@

We thank the anonymous referee, whose input has significantly enhanced the quality of this work, particularly for giving up their time to review the work during a unprecedented time of global crisis. We thank Maryam Modjaz and Dan Perley for providing

@@

@@

%\newpage
\section{Bayesian Fits - sSFR}
\label{app:b}

\begin{figure*}
\includegraphics[width=0.45\textwidth]{mc_figs/rets_s12_histfit_ssfr_final.png}
\caption{Results of the MCMC fits to the PDFs of sSFR for DES RETs and SESNe, accounting for uncertainties in each bin.
\label{fig:histfit_s12_ssfr}}
\end{figure*}

\begin{figure*}
\includegraphics[width=\textwidth]{mc_figs/rets_s12_param_diffs_ssfr_final.png}
\caption{Histograms showing the differences between the fit parameters across the MCMC samples for the comparison between RETs and SESNe. 
\label{fig:param_diffs}}
\end{figure*}

>>>>

% Abstract of the paper
\begin{abstract}
Rapidly evolving transients (RETs), also termed fast blue optical transients (FBOTs), \replyref{are a recently-discovered group of astrophysical events that display rapid luminosity evolution. RETs typically rise to peak in less than 10 days and fade within 30, a timescale unlikely to be compatible with the decay of Nickel-56 that drives conventional SNe. Their peak luminosity spans a range of $-15 <M_g<-22.5$, with some events observed at redshifts greater than 1.} Their evolution on fast timescales has hindered high quality follow-up observations, and thus their origin and explosion/emission mechanism remains unexplained. In this paper we \replyref{present} the largest sample of RETs to date, comprising 106 objects \replyref{discovered by} the Dark Energy Survey, and perform the most comprehensive analysis of RET host galaxies. Using deep-stacked photometry and emission-lines from OzDES spectroscopy, we derive stellar masses and star-formation rates (SFRs) for 49 host galaxies, and metallicities \replyref{([O/H])} for 37. We find that RETs explode exclusively in star-forming galaxies and are thus likely associated with massive stars. Comparing RET hosts to samples of host galaxies of other explosive transients as well as field galaxies, we find that RETs prefer galaxies with high specific SFRs \replyref{($\langle\log \left(\mathrm{sSFR}\right)\rangle \sim -9.6$ ),} indicating a link to young stellar populations, similar to stripped-envelope SNe. RET hosts appear to show a lack of chemical enrichment, their metallicities akin to long duration gamma-ray bursts and superluminous SN host galaxies \replyref{($\langle12+\log\left(\mathrm{O/H}\right) \rangle \sim 9.4$)}. There are no clear relationships between \replyref{mass or SFR} of the host galaxies and the peak magnitudes or decline rates of the transients themselves.

@@

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
\replyref{}
\section{Introduction}

@@

There are a small number of other objects in the literature that would pass the DES RET criteria, including SNLS04D4ec \citep{Arcavi2016} at $z\sim0.59$, \replyref{the radio-bright} ZTF18abvkwla \replyref{ at $z=0.27$. whose rapid evolution and radio brightness have been attributed to a shock and a collimated jet, respectively \citep{Ho2020}, and the rapidly-evolving SN Ic-BL-like iPTF16asu \citep{Whitesides2017} at $z\sim0.19$. }

There are also a limited number of events detected in the local Universe whose properties appear consistent with the RETs seen at cosmological distances in the PS1 and DES samples, the most widely studied of which is AT2018cow \citep[e.g.][]{Prentice2018,Margutti2019,Perley2019}. \replyref{AT2018cow has drawn many comparisons to the cosmological RETs from DES and PS1 \citep[e.g.][]{Perley2019,Margutti2019,Fox2019,Mohan2020} due to its rapid evolution and blue colour. The brightness of AT2018cow declined from its discovery, with constraints on the rise time of 1 day, and from X-rays through to radio wavelengths did not resemble any known SN, GRB afterglow, or kilonova (KN; \citealt{Ho2019}). It displayed a contracting photosphere as well as evidence for central-engine power alongside an unusual spectrum that showed similarities to broad-lined SNe Ic (SN Ic-bl) at early stages \citep[e.g.][]{Xu2018,Izzo2018}, developing to something entirely different at later epochs \citep{Perley2019} with hints of similarities to interacting SNe Ibn \citep{Fox2019}.} There are many diverse explanations for the power source of AT2018cow \replyref{proposed} in the literature, including: \replyref{engine-driven stellar explosions \citep{Ho2019,Margutti2019,Mohan2020}}; electron capture collapse of merged white dwarfs \citep{Lyutikov2019}; a tidal disruption event (TDE) of a white dwarf \citep{Kuin2019} or of a main sequence star by an intermediate mass black hole \citep{Perley2019}; common envelope jets supernova (CEJSN; \citealt{Soker2019}); or a wind-driven transient \citep{Piro2020,Uno2020}.
Other nearby \replyref{($z<0.1$)} rapid transients \replyref{that would pass the DES search criteria} include the local fast-declining SN-like transient SN2018kzr \citep{McBrien2019} which is explained by the accretion-induced collapse of a white dwarf or a white dwarf-neutron star/\replyref{black hole} merger \citep{Gillanders2020}, KSN-2015K \citep{Rest2018} whose fast rise and decline is explained by the shock of an SN running into previously-expelled material, \replyref{SN2018gep \citep{Ho2019a} \replyref{which is a rapidly-declining, spectroscopically classified, nearby SN Ic-BL with a rapid rise to peak}, and the H-rich, relativistic CSS161010 \citep{Coppejans2020}.} \replyref{These objects evolved rapidly enough to have passed the DES RET criteria (Section \ref{subsec:new_method}) even at $z\sim1$ where they would appear slower, although they (or their host galaxies) may not necessarily have been bright enough to be detected.}

@@

The original method of finding RETs in the DES-SN data (and presented in \citetalias{Pursiainen2018}) was designed to be simple and used light curve modelling with Gaussian and linear fits. The simplistic method made it possible to look for exotic transients without knowing their observed characteristics beforehand and resulted in a large sample of photometrically selected fast transients. However, as the search was simplistic and relied heavily on visual inspection of the available data (etc. images, light curves, host galaxy information), it is impossible to quantify the completeness of the sample. For instance, due to the large redshift range within the sample it is entirely possible that distant events with longer rise times could have been missed due to time dilation stretching their lightcurves, while faster events \replyref{at} low redshift may evolve on timescales quicker than the survey cadence. Here, a more sophisticated search method is presented. As only a fraction of transients in DES-SN have redshift information from their host galaxies, we perform the search in the observer frame. The key features of RETs that separate them from most traditional SNe types are the fast light curve evolution \replyref{($t_{\mathrm{half}}\leq12$d)} and blue colour at peak ($g-r \lesssim 0.5$). Even though both of these quantities depend on the redshift of the transient \replyref{(timescales, for example, are dilated by a factor of 2 at $z=1$ while higher-redshift objects may appear bluer, since the hot and bright UV is shifted into the observed bands)}, they still effectively distinguish the fast events from traditional SN types. \replyref{Here, we} select a sample based on observed rise times and colours at peak brightness.

@@

In order to apply selection criterion i) from Section \ref{subsubsec:trise_gr} it is necessary to distinguish between DES-SN candidates that are truly multi-season events (typically active galactic nuclei; AGN) and those that are single-season events with spurious detections in other seasons. 
To detect and filter out AGN we use a basic convolutional neural network (CNN) classifier. \replyref{The CNN uses features of the GP light curves in a set of layers in order to ``learn" the difference between SN and AGN lightcurves. We implement two 1-dimensional feature layers: the first searches for evolution over 2-day timescales in all four filters, while the second deals with long-term evolution on 20-day timescales. We include a dropout layer after each filter layer to prevent over-fitting, and a max-pooling layer to increase speed. At the end, a softmax layer outputs a probability that the individual transient is SN-like, for which we use 50\% as the threshold}.  We train the CNN on spectroscopically confirmed SNe of all types as well as the 92 RETs identified using the \citetalias{Pursiainen2018} method (181 objects) and on spectroscopically typed AGN (182 objects), and use it to separate the sample into two photometric subtypes: AGN-like and SN-like. The classifier returns SNe-like objects with an accuracy of 0.992 on the test set (\replyref{ the remaining 391 DES SNe and 79 AGN that were spectroscopically classified at the time of testing}). No SNe were classified as AGN-like. The remaining AGN classified as SNe-like are removed by manual vetting later in the process. \replyref{While this performance is very encouraging, it should be noted that the training and validation being performed on spectroscopically typed objects, which are in general high SNR, means its performance on fainter/lower-SNR objects is not well known.}  The CNN does not separate the SN-like objects into RETs and other SN subtypes: this is done in the subsequent processing step.

@@

To show how RETs compare to the galaxy population as a whole, we use a sample of $\sim800,000$ measurements from the MPA-JHU catalogues of stellar masses (based on the methods of \citealt{Kauffmann2003,Salim2007}), SFRs (based off \citealt{Brinchmann2004}), and metallicities (based off \citealt{Tremonti2004}) from the  catalogues of SDSS Data Relase 7 (DR7; \citealt{Abazajian2009}). The mean redshift is 0.08 which is much lower than for the RETs, such that significant evolution in the galaxy population has happened between the majority of RET hosts and the SDSS sample. We do not correct for this, but take it into account when analysing our findings. \replyref{The magnitude-limited nature of the SDSS redshift catalogue means that it is biased against low-mass galaxies, and metallicity estimates requiring strong lines to pass SNR thresholds mean that they too will be biased. Many works exploring extreme transients such as LGRBs, GRB-SNe, and SLSNe, use a complete, volume limited sample of field galaxies such as the Local Volume Legacy Survey (LVL) sample as a comparison \citep[e.g.][]{Chen2017,Modjaz2019,Taggart2019}. However the spectroscopic follow-up of DES RETs was performed as part of the magnitude-limited OzDES survey (Section \ref{subsec:spec}), which has similar selection biases and incompleteness effects to SDSS, albeit shifted to a higher redshift. In comparison, the many dwarf galaxies in LVL would not be detected in DES photometry, or get a redshift measurement from OzDES, and thus such a comparison sample is not representative of the way the DES RET host sample was observed. For completeness and illustrative purposes only, we display both the SDSS and LVL (data from D. Perley, priv. communication) comparison samples.}

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{figs/RET_BPT.png}
\caption{Baldwin-Phillips-Terlovich (BPT) diagram for RET hosts, showing that the emission lines are consistent with being generated by star formation rather than AGN activity. The three curves show the delimitation between star formation and AGN at $z =0, 0.5, 1$ according to \citet{Kewley2013}. The redshift of each RET is indicated by the colourbar on the right.
\label{fig:bpt}}
\end{figure}
\begin{table*}

@@

Due to the low S/N of the spectra in this sample, we are constrained to a subset of metallicity diagnostics by the availability of only a handful of the strongest emission lines, namely \halpha, \hbeta, \OII 3727, \OIII 4959/5007, \NII 6548/6583, and \SII 6717/6731. Furthermore, for each host galaxy only a subset of these lines is detected - for example, \halpha, \NII~and \SII~are redshifted out of the spectral coverage at $z>0.3$, leaving only the oxygen and \hbeta~lines available and thus the R23 diagnostic, which is based off the relative strengths of the \OII and \OIII lines. For hosts at $z<0.3$ we are able to use the \OIII /\NII~(O3N2), \NII /\halpha~(N2), and \SII /\NII~(S2N2) line ratios. 
\begin{figure}

\includegraphics[width=0.5\textwidth]{figs/SFR_Mike_RETs.png}
\caption{The $M_* -$ SFR sequence of RET hosts, hosts of other rapid transients and SNe II, as well as SDSS field galaxies (grey contours) and \replyref{LVL galaxies (grey circles)}. SFRs have not been corrected for redshift evolution. RET hosts lie slightly above the low-$z$ star-formation main sequence upper half of the SDSS contours, and systematically avoid passive galaxies (dense SDSS contours in the lower right).
\label{fig:sfms_sfr}}
\end{figure}
Due to the redshift range of our sample, and the limited wavelength coverage of the spectra ($3000-8000$~\AA), we are unable to use a single line ratio to estimate the oxygen abundances. We thus determine a set of indicators for which to calculate abundances. For the O3N2 and N2 indicators we use the calibration of \citet{Pettini2004} (PP04), and if \SII~is detected we derive an abundance using the S2N2 diagnostic of \citet{Dopita2016} (D16). For the R23 indicator, we use the calibration of \citet{Kobulnicky2004} (KK04). At abundances around $12 + \log \mathrm{(O/H)} \sim 8.4$, the R23 indicator becomes two-tailed, with a low and a high value of metallicity corresponding to a single R23 ratio. In cases where the lines are available, we break this degeneracy by cross-calibrating with the \NII/\OII~ratio \citep{Kewley2008}. In the cases where \NII~is not available, and there are no other diagnostics that can be used to inform the choice of branch, we use the host galaxy stellar mass to derive a crude metallicity estimate from the mass-metallicity relation (MZR) of \citet{Kewley2008} based upon the PP04 O3N2 diagnostic. For $12 + \log \mathrm{(O/H)}_{\mathrm{MZR}} < 8.4$ we chose the lower branch, while for higher MZR metallicities we choose the upper branch. We note that this is a rough estimation. If we leave the branch choice for those with no \NII~to be random, we find that the results are consistent to within uncertainties.

The samples to which we compare metallicities span different redshift ranges, were observed with different equipment, and in many cases were compiled before certain (particularly the D16) diagnostics were devised. Therefore, in order to compare oxygen abundances between different samples we transform all abundances onto the PP04 O3N2 scale using the conversion factors given in \citet{Kewley2008}. This is not possible for the D16 diagnostic, so we discard it from the rest of our analysis, although for completeness we provide it for DES RET hosts where available. For samples that quoted multiple diagnostics, or for which sufficient line flux measurements were provided from which to calculate multiple diagnostics, we transform them all to the PP04 O3N2 scale. Following the prescription of \citet{Kruehler2015} we simultaneously minimise the oxygen abundance against the PDFs of the various different diagnostics scaled to PP04 O3N2, resulting in a final `best' PDF. We take $1\sigma$ uncertainties from the 16\textsuperscript{th} and 84\textsuperscript{th} percentiles of these PDFs, and for DES RETs display the results in Table \ref{tab:derived}.

@@

We thank the anonymous referee, whose input has significantly enhanced the quality of this work, particularly for dedicating their time to review the work during unprecedented times of global crisis. We thank Maryam Modjaz and Dan Perley for providing the LVL compilation data.

@@

\section*{Data availability}
The data underlying this article are available in the article and in its online supplementary material.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%\newpage
\section{Bayesian Fits - sSFR}
\label{app:b}
\begin{figure*}
\includegraphics[width=\textwidth]{mc_figs/rets_s12_param_diffs_ssfr_final.png}
\caption{Histograms showing the differences between the fit parameters across the MCMC samples for the comparison between RETs and SESNe. 
\label{fig:param_diffs}}
\end{figure*}
